---
title: "House price regression"
author: "Bruno Pilarczyk Tanguy Magon"
date: "20/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
graphics.off()
```


```{r libraries, include=FALSE}
if (require(data.table) == FALSE)
  install.packages("data.table")
library("data.table")

if (require(xgboost) == FALSE)
  install.packages("xgboost")
library("xgboost")

if (require(tidyverse) == FALSE)
  install.packages("tidyverse")
library("tidyverse")

if (require(BBmisc) == FALSE)
  install.packages("BBmisc")
library("BBmisc")

if (require(corrplot) == FALSE)
  install.packages("corrplot")
library("corrplot")

if (require(e1071) == FALSE)
  install.packages("e1071")
library("e1071")

if (require(plyr) == FALSE)
  install.packages("plyr")
library(plyr)

if (require(pROC) == FALSE)
  install.packages("pROC")
library(pROC)

if (require(fastDummies) == FALSE)
  install.packages("fastDummies")
library(fastDummies)

if (require(kernlab) == FALSE)
  install.packages("kernlab")
library(kernlab)

if (require(DMwR) == FALSE)
  install.packages("DMwR")
library(DMwR)

if (require(grid) == FALSE)
  install.packages("grid")
library("grid")

if (require(gridExtra) == FALSE)
  install.packages("gridExtra")
library("gridExtra")

if (require(caret) == FALSE)
  install.packages("caret")
library("caret")

if (require(rpart) == FALSE)
  install.packages("rpart")
library("rpart")

if (require(AppliedPredictiveModeling) == FALSE)
  install.packages("AppliedPredictiveModeling")
library("AppliedPredictiveModeling")

if (require(RColorBrewer) == FALSE)
  install.packages("RColorBrewer")
library("RColorBrewer")

if (require(pls) == FALSE)
  install.packages("pls")
library("pls")
```

#Data import
```{r 1 path name}
#Path of the files:
path = "/Users/brunopilarczyk/Documents/Cours/M2 Stat Eco/S1/Mathematics of deep learning algo./Machine Learning/Project-files/Regression/train.csv"

#Importing both datasets:
houses = data.frame(fread(file = path, dec = ".", header = T))
```

```{r 1 creating a test set}
## 80% of the sample size
smp_size = floor(0.8 * nrow(houses))


set.seed(123)
train_ind = sample(seq_len(nrow(houses)), size = smp_size)

train <- houses[train_ind, ]
train$test = 0
test <- houses[-train_ind, ]
test$test = 1

houses = rbind(train, test)
```

```{r Dataset structure}
str(houses)
sum(is.na(houses))
```

There are a lot of string variables which needs to be cleaned and transformed in order to be handled. Furtheremore we need to take care of NAs.

#1 Data Cleaning
```{r 1 Missing values per variables}
apply(houses, 2, function(x) sum(is.na(x)))
```

We can erase the variables containing too much NAs.

```{r 1 Deleting variables with loads of NAs}
unuseful_col = c("Alley", "FireplaceQu", "PoolQC", "Fence", "MiscFeature")

houses = houses %>% select(-unuseful_col)
```
Median imputation for missing values concerning numerical variables:

```{r 1 median imputation}
list1 = c("LotFrontage", "GarageYrBlt", "MasVnrArea")

for (word in list1){
  houses[[word]] = houses[[word]] %>%
    replace_na(median(houses[[word]][is.na(houses[[word]]) == FALSE]))
}
```

Now we have to take care of missing values regarding cateogrical variables, first let us transform them into factor:

```{r 1 categorical columns}
variables = colnames(houses)
vec1 = numeric(length(houses))
for (i in 1:length(houses)){
  vec1[i] = is.numeric(houses[[variables[i]]])
}

columns_cat = colnames(houses[,which(vec1== 0)])
```

```{r 1 string variables factorization}
for (word in columns_cat){
  houses[[word]] = factor(houses[[word]])
}
houses = knnImputation(houses, k = 5, meth = "median")
```

In order to create dummies for each of each categorical variables, we can apply the following double-loop.

Firstly, it creates the list of the modalities for a given variables, then it creates a dummy for each modality.

Afterwards, it erases the last modality creates in order to avoid collinearity between variables.

Lastly, it deletes the original categorical variable such that only its modalities transformed into dummies remain.

```{r 1 double-loop for creating dummies}
for (variable in columns_cat){
  list_modalities = unique(as.character(houses[[variable]]))
  for (word in list_modalities){
    houses[[paste(word,"_",variable, sep ="")]] = ifelse(houses[[variable]] == word, 1, 0)
  }
  houses = houses[,-length(houses)]
  houses = houses %>% select(-variable)
} 
```

Data cleaning is over.

#2 Data exploration

```{r 2 Deleting ID column}
houses = houses[,-1]
```


```{r 2 summary}
summary(houses)
```

```{r 2 Descriptive statistics}
qplot(x = houses[,"SalePrice"], fill= ..count.., geom = "histogram", bins = 30) +
  xlab("Sale Price") +
  ylab("Frequency") +
  scale_fill_gradient(low = "grey", high = "blue") +
  theme_light(base_family = "Times") +
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size = 12)) 
```

```{r 2 PCA}
# The 37th and 38th columns are respectively the SalePrice (target variable) and the 
acp=prcomp(houses[,-c(37,38)])
plot(acp)
biplot(acp)
plot(acp$x)
```

#3 Data preparation
We can split the dataset into both train set and test set thanks to the variable test we created before. We erase this variable in the mean time.
```{r 3 Splitting houses into train and test}
train = houses[houses$test == 0, -38]
test = houses[houses$test == 1, -38]
```

In order to do cross validation, we create 4 segments of the data set. This resampling procedures insures we will not overfit to the original train set.

#4 Model

```{r 4-fold cross validation}
k_fold_cv = trainControl("cv", number = 4)
```

#4.1 Ridge regression

Let's begin by the Ridge regression in order to find a good set of parameters. We set alpha to 0 in order to do Ridge regression. We will minimize this loss function:

$$\displaystyle \sum_{i = 1}^n||y_i - <x_i,\beta>||_2^2 + \lambda||\beta||_2^2$$ 

Furtheremore we will consider centered and scaled data such as we will remove to each variables its mean and divide this difference by the standard deviation of the variable as followed:

$$\dfrac{x_i - \overline{x}}{\sigma_x}$$

```{r 4.1 Ridge regression}
# Build the model
set.seed(1)

lambda_ridge <- 10^seq(0, 5, length = 100)

ridge = train(
  SalePrice ~., data = train,
  method = "glmnet",
  trControl = k_fold_cv,
  tuneGrid = expand.grid(alpha = 0, lambda = lambda_ridge),
  preProcess = c("center", "scale")
  )
#Coefficients of the best model
coeff_ridge = coef(ridge$finalModel, ridge$bestTune$lambda)

# Make predictions
pred_train_ridge = ridge %>% predict(train)
pred_test_ridge = ridge %>% predict(test)
# Model prediction performance
ridge_result = data.frame(
  RMSE_train = RMSE(pred_train_ridge, train$SalePrice),
  RMSE_test = RMSE(pred_test_ridge, test$SalePrice)
)

regul_param_ridge = ggplot() +
  geom_point(aes(x = ridge$results$lambda, y = ridge$results$RMSE)) +
  geom_line(aes(x = ridge$results$lambda, y = ridge$results$RMSE), colour = "blue")  +
  xlab("Regularization parameter") +
  ylab("Average RMSE on 4-folds") +
  theme_light(base_family = "Times")
regul_param_ridge

residuals_ridge = ggplot() +
  geom_point(aes(x = ridge %>% predict(test), y = test$SalePrice - ridge %>% predict(test))) +
  geom_line(aes(x = ridge %>% predict(test), y = 0), colour = "purple") +
  xlab("Fitted values") +
  ylab("Residuals") +
  theme_light(base_family = "Times")
residuals_ridge
ridge_result
```
We can also find the optimal value for $\lamdda$ through cross-validation:

```{r 4.1 Optimal regularization parameter}
opt_lambda_ridge = ridge$bestTune$lambda
opt_lambda_ridge
```

`opt_lambda_ridge` is the value that minimizes the RMSE for the Ridge regression model.

#4.2 LASSO regression

In order to set the norm associated to the parameters in the penalty part of the loss function, we set alpha to 1 in the paramrters of the train function. Here is the loss function to minimize:

$$\displaystyle \sum_{i = 1}^n||y_i - <x_i,\beta>||_2^2 + \lambda||\beta||_1$$ 

```{r 4.2 Lasso}  
# Build the model
lambda_lasso <- 10^seq(0, 4, length = 100)

lasso = train(
  SalePrice ~., data = train,
  method = "glmnet",
  trControl = k_fold_cv,
  tuneGrid = expand.grid(alpha = 1, lambda = lambda_lasso),
  preProcess = c("center", "scale")
  )
#Coefficients of the best model
coeff_lasso = coef(lasso$finalModel, lasso$bestTune$lambda)

# Make predictions
pred_train_lasso = lasso %>% predict(train)
pred_test_lasso = lasso %>% predict(test)
# Model prediction performance
lasso_result = data.frame(
  RMSE_train = RMSE(pred_train_lasso, train$SalePrice),
  RMSE_test = RMSE(pred_test_lasso, test$SalePrice)
)

regul_param_lasso = ggplot() +
  geom_point(aes(x = lasso$results$lambda, y = lasso$results$RMSE)) +
  geom_line(aes(x = lasso$results$lambda, y = lasso$results$RMSE), colour = "red")  +
  xlab("Regularization parameter") +
  ylab("Average RMSE on 4-folds") +
  theme_light(base_family = "Times")
regul_param_lasso

residuals_lasso = ggplot() +
  geom_point(aes(x = lasso %>% predict(test), y = test$SalePrice - lasso %>% predict(test))) +
  geom_line(aes(x = lasso %>% predict(test), y = 0), colour = "purple") +
  xlab("Fitted values") +
  ylab("Residuals") +
  theme_light(base_family = "Times")
residuals_lasso
lasso_result
```
We can also find the optimal value for $\lamdda$ through cross-validation:

```{r 4.2 Optimal regularization parameter}
opt_lambda_lasso = lasso$bestTune$lambda
opt_lambda_lasso
```

`opt_lambda_lasso` is the value that minimizes the RMSE for the LASSO regression model.

```{r 4.3 Model comparison}
models <- list(ridge = ridge, lasso = lasso)
resamples(models) %>% summary( metric = "RMSE")

ridge_result
lasso_result
diff = lasso_result - ridge_result
diff$RMSE_test
```


Even if Ridge is better on the 4-folds and on the whole train set, LASSO model works better on the test set by decreasing the RMSE by `diff$RMSE_test`.

```{r 4.3 coefficients}
head(coeff_ridge, 20)
head(coeff_lasso, 20)
```

As expected, some coeffcients are exactly equal to 0 in the LASSO regression because of the penalty term which is the regularization parameter times the L1 norm of the coefficients.
